---
title: "Literacy population of Age group (15-24)"
author: "Binod Jung Bogati"
date: "7/19/2018"
output: 
  html_document:
    keep_md: yes
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```


The data set is provided by Open Nepal which contains literacy population of age group (15-24) based on zone, region, districts.

# Import the data file

```{r}
literacy <- read.csv("./data/literacy-rate.csv", stringsAsFactors = FALSE)

head(literacy)
```



# Rename the column headers

```{r}
names(literacy) <- c("district", "zone", "geo_region", "dev_region", "group", "gender", "population")

head(literacy)
```



# Rename the population sub-groups

```{r}
# get number of rows
nrow(literacy)

# renaming is needed for any long names
pop <- lapply(literacy$group, function(x) if(x == "Literate Population by age 15 - 24 "){ x = "literate_pop"} else{ x = "total_pop" } )
literacy$group <- sapply(pop, paste0, collapse="") 

# cleaning gender column

library(stringr)
library(dplyr)

replace_sp <- "[[:space:]].*"

literacy <- literacy %>% 
  mutate(gender = str_replace_all(gender, replace_sp, "")) 

```

# Get the literate population

```{r}
literacy_pop <- literacy %>% 
  filter(group == "literate_pop")
```

```{r}
library(ggplot2)

ggplot(literacy_pop, aes(x = gender, y = population)) +
  geom_bar(stat = "identity", aes(fill = zone), position = "dodge") +
  ggtitle("Literate Population of Age (15 - 24) by Zones")

```



# Literate population by gender

```{r}
get_population <- literacy %>% 
  filter(group == "total_pop")

head(get_population)
```

```{r}
total_male <- get_population %>% 
  filter(gender == "Male")

total_male_pop <- sum(total_male$population)

total_male_pop

```
```{r}
total_female <- get_population %>% 
  filter(gender == "Female")

total_female_pop <- sum(total_female$population)
total_female_pop
```

# Total population of age group 15-24

```{r}
total_population <- total_male_pop + total_female_pop
total_population
```

# Total literate population

```{r}

head(literacy_pop)

literate_male <- literacy_pop %>% 
  filter(gender == "Male")

head(literate_male)

total_literate_male <- sum(literate_male$population)
total_literate_male
```

```{r}
literate_female <- literacy_pop %>% 
  filter(gender == "Female")

total_literate_female <- sum(literate_female$population)
total_literate_female
```

# Total literate population of age group (15-24)

```{r}
total_literate_population <- total_literate_male + total_literate_female
total_literate_population
```

# Total illiterate population of age group (15-24)

```{r}
illiterate_population <- total_population - total_literate_population
illiterate_population
```

# Creating a dataframe

```{r}
population_summary <- data.frame(
  c("literate_pop", "literate_pop", "total_pop", "total_pop"),
  c("male", "female", "male", "female"), 
  c(2234745, 2246814, 2487172, 2802879), stringsAsFactors = FALSE)

names(population_summary) <- c("group", "gender", "population")

head(population_summary)
```

```{r}

literate_gender <- population_summary %>% 
  filter(group == "literate_pop")

literate_gender$percent[1] <- literate_gender$population[1] / sum(literate_gender$population) * 100
literate_gender$percent[2] <-  literate_gender$population[2] / sum(literate_gender$population) * 100

ggplot(literate_gender, aes(x = "", y = population, fill = gender)) +
  geom_bar(stat = "identity", width = 0.5) +
  coord_polar("y", start=0) +
  labs(title="Literate Population by Gender (Age 15-24)", caption= "Male : 49.87 % \n Female : 50.13 %") +
  geom_text(aes(label = population), position = position_stack(vjust = 0.5), size = 3) 
```


```{r}
eastern_region <- literacy %>% 
  filter(dev_region == "Eastern")

literate_east <- eastern_region %>% 
  filter(group == "literate_pop")

total_east <- eastern_region %>% 
  filter(group == "total_pop")

```

```{r}
ggplot(literate_east, aes(x = gender, y = population)) +
  geom_bar(stat = "identity", aes(fill = zone), position = "dodge") +
  ggtitle("Literacy population (Age: 15-24) of Eastern Region")

ggplot(total_east, aes(x = gender, y = population)) +
  geom_bar(stat = "identity", aes(fill = zone), position = "dodge") +
  ggtitle("Total population (Age: 15-24) of Eastern Region")
  
```

```{r}
literacy_sagarmatha <- literate_east %>% 
  filter(zone == "Sagarmatha")

total_sagarmatha <- total_east %>% 
  filter(zone == "Sagarmatha")

```

```{r}

ggplot(literacy_sagarmatha, aes(x = district, y = population)) +
  geom_bar(stat = "identity", aes(fill = gender), position = "dodge") +
  ggtitle("Literate population of Sagarmatha zone (Age 15-24)")

ggplot(total_sagarmatha, aes(x = district, y = population)) +
  geom_bar(stat = "identity", aes(fill = gender), position = "dodge") +
  ggtitle("Total population of Sagarmatha zone (Age 15-24)")
```

